Analysis

### 2023-03-27 Analysis.rmd
restructure code file

@@

# Define Helper Functions, Load Data

@@

# Figure SI.1: Data and data revisions
data revisions: https://github.com/CSSEGISandData/COVID-19/issues/6021
https://coronavirus.data.gov.uk/details/whats-new 
https://coronavirus.data.gov.uk/details/whats-new/record/965a7923-dacf-42ed-abb9-3a972dbf1835

# Figure SI.4: Weekly participation humans and hub ensemble

## Performance

@@

ggsave("output/figures/performance.png", plot = p_performance, 
       width = 7, height = 6)

scores$model |> unique()

forecasts_comparison_revised <- filter_comparison_rt_direct(forecasts_revised)
scores_forecasts_comparison_revised <- score_and_add_log(forecasts_comparison_revised)

scores_forecasts_comparison |> 
  filter(horizon == 2) |>
  mutate(name = gsub("\ \\(Rt\\)", "", model)) |>
  summarise_scores(by = c("target_type", "type_and_scale", "name", "method", "scale")) |>
  ggplot(aes(y = interval_score, x = name, group = method, color = method)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~ type_and_scale, scale = "free_y") + 
  theme_scoringutils()

ggsave("output/figures/comparison-direct-rt-individual.png", 
       plot = p_comparison_direct_rt_individual, 
       width = 7, height = 4)

p_comparison_direct_rt_individual_reivsed <- plot_rt_direct_individual(scores_forecasts_comparison_revised)

ggsave("output/figures/comparison-direct-rt-individual-revised.png", 
       plot = p_comparison_direct_rt_individual, 
       width = 7, height = 5)

# performance plot for an ensemble of the subset of human and direct forecasts
forecasts_comparable_ensemble <- forecasts_comparison |>
  group_by(target_type, horizon, method, quantile, true_value, forecast_date, target_end_date) |>
  summarise(prediction = median(prediction)) |>
  mutate(model = str_to_title(method)) |>
  as.data.table()

forecasts_comparable_ensemble_revised <- forecasts_comparison_revised |>
  group_by(target_type, horizon, method, quantile, true_value, forecast_date, target_end_date) |>
  summarise(prediction = median(prediction)) |>
  mutate(model = str_to_title(method)) |>
  as.data.table()

scores_comparable_ensemble_revised <- score_and_add_log(forecasts_comparable_ensemble_revised)

p_performance_comparison_revised <- plot_performance(
  scores_comparable_ensemble_revised
)

ggsave("output/figures/performance-rt-comparison-revised.png", plot = p_performance_comparison, 
       width = 7, height = 8)

scores_comparable_ensemble |>
  filter(horizon == 2) |>
  summarise_scores(by = c("target_type", "horizon", "model", "scale"))

scores |>
  filter(scale != "sqrt") |>
  filter(model %in% models_non_human) |>
  filter(horizon == 2,
         scale == "log") |>
  summarise_scores(by = c("target_type", "horizon", "model", "scale"))

```{r}
scores_summarised <- scores |>
  filter(scale %in% c("log", "natural")) |>
  filter(model %in% models_non_human) |>
  summarise_scores(by = c("model", "target_type", "scale", "horizon", "forecast_date")) |>
  group_by(target_type, scale, horizon) |>
  # mutate(interval_score = (interval_score - mean(interval_score)) / sd(interval_score)) |>
  group_by(model, target_type, scale, horizon) |>
  summarise(mean = mean(interval_score), 
            sd = sd(interval_score)) |>
  pivot_wider(id_cols = c(model, target_type, horizon), 
              names_from = scale, values_from = c(mean, sd), names_prefix = "stand_") |>
  arrange(target_type, model)
```

# Sinificance and bootstrap analysis

```{r}
scores |>
  filter(model %in% models_non_human) |>
  filter(horizon == 2) |>
  summarise_scores() |>
  select(model, forecast_date, type_and_scale, interval_score) |>
  pivot_wider(names_from = model, values_from = interval_score) |>
  group_by(type_and_scale) |>
  summarise(p1 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-direct`, paired = TRUE)$p.value, 
            p2 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-ensemble`, paired = TRUE)$p.value, 
            p3 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-rt`, paired = TRUE)$p.value)

wilcox.test(1:20, 2:21) |> str()

Table with performance

```{r, performance-table}
make_performance_table <- function(
    scores, scores_summarised, 
    hor = 2, 
    cap = "Performance for two-week-ahead forecasts \\label{tab:scores}"
) {
  scores |>
    filter(scale %in% c("log", "natural")) |>
    summarise_scores(by = c("model", "target_type", "scale", "horizon")) |>
    select(model, target_type, scale, interval_score, coverage_50, coverage_90, horizon) |>
    pivot_wider(names_from = scale, values_from = interval_score) |>
    select(model, target_type, natural, log, coverage_50, coverage_90, horizon) |>
    inner_join(scores_summarised) |>
    mutate(across(c(where(is.numeric)), \(x) signif(x, digits = 2))) |>
    arrange(target_type, model) |>
    mutate(natural = label_fn(natural), 
           log = label_fn(log)) |>
    mutate(natural = paste0(natural, " (", label_fn(sd_stand_natural), ")"), 
           log = paste0(log, " (", label_fn(sd_stand_log), ")")) |>
    filter(horizon == hor) |>
    select(-mean_stand_log, -sd_stand_log, -mean_stand_natural, -sd_stand_natural, -horizon) |>
    kable(format = "latex", booktabs = TRUE, 
          align = c("l", "l", rep("c", 4)),
          caption = cap,
          linesep = c('', '', '', '\\addlinespace'),
          col.names = c("Model", "Target", "WIS", "WIS (log scale)", 
                        "Coverage 50%", "Coverage 90%")) |> 
    kable_styling(latex_options = c("scale_down", 
                                    "hold_position")) 
}

@@

ggsave("output/figures/scores-and-forecasts.png", plot = p_scores_and_pred, 
       width = 9, height = 9)

## Look at expert forecasts

# ensemble of experts and non-experts

create_expert_ensemble <- function(forecasts) {

@@

expert_ensemble$scores |>
   make_performance_table(
     scores_summarised = expert_scores_summarised,
    cap = "Performance for four-week-ahead forecasts of experts and non-experts. Standard deviations are given in brackets. \\label{tab:scores-experts-4}"
    ) %>%
  gsub("\\{table\\}", "\\{table\\*\\}", x = .) |>
  cat(file = 'manuscript/performance-table-experts-4.tex')

@@

expert_ensemble_direct <- forecasts |>
  filter(model %in% models_direct) |>
  create_expert_ensemble()

expert_ensemble_direct$scores |>
   make_performance_table()

p_scores_expert_direct <- plot_performance(expert_ensemble_direct$scores)
ggsave("output/figures/performance-expert-direct.png", 
       plot = p_scores_expert_direct, 
       width = 7, height = 6)

# plot_scores_and_pred(
#   forecasts_expert_ensemble,
#   scores_expert_ensemble
# )

@@

```{r}
scores_df <- scores  |>
  update_expert_status() |>
  mutate(expert = ifelse(expert, "Expert", "Non-Expert")) |>
  filter(model %in% models_human) |>
  filter(horizon == 2) |>
  summarise_scores()

mean_wis <- scores_df |>
  group_by(scale, expert, target_type, type_and_scale) |>
  summarise(wis_mean = signif(median(interval_score), 3)) |>
  mutate(wis_mean = paste("Mean:", label_fn(wis_mean)))

scores_df |>   
  ggplot(aes(x = expert, y = interval_score)) +
  geom_violin(alpha = 0.2) + 
  geom_boxplot(alpha = 0.5) + 
  # scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  geom_text(data = mean_wis, 
            aes(label = unique(wis_mean), y = -Inf), 
            vjust   = -1) + 
  facet_wrap(~ type_and_scale, scale = "free_y") + 
  scale_y_continuous(labels = label_fn, 
                     expand = expansion(mult = c(0.15, 0.05))) + 
  theme_scoringutils()

scores_df |>
  summarise_scores(by = c("expert", "target_type", "scale", "horizon")) |>
  rename(model = expert) |>
  make_performance_table()

@@

```{r}
forecasts$forecast_date |>
  unique()

forecasts$target_end_date |>
  unique()
```

Interesting questions
- [x] performance for cases and deaths
- [x] on the natural scale and log scale

- performance experts vs. non-experts
- performance Rt vs. non-Rt forecast
- need a table with number of participants
- add calibration plot (maybe just copy the plot from the Germany / Poland paper)
- need to find the numbers of member models

Methods
- Hub ensemble switched from mean to median on 2021-07-19

>>>>

# Setup, define Helper Functions, Load Data
================================================================================

@@

# Helper functions
# ==============================================================================

@@

# Table 1: Performance
================================================================================

```{r, performance-table}
scores_summarised <- scores |>
  filter(scale %in% c("log", "natural")) |>
  filter(model %in% models_non_human) |>
  summarise_scores(by = c("model", "target_type", "scale", "horizon", "forecast_date")) |>
  group_by(target_type, scale, horizon) |>
  # mutate(interval_score = (interval_score - mean(interval_score)) / sd(interval_score)) |>
  group_by(model, target_type, scale, horizon) |>
  summarise(mean = mean(interval_score), 
            sd = sd(interval_score)) |>
  pivot_wider(id_cols = c(model, target_type, horizon), 
              names_from = scale, values_from = c(mean, sd), names_prefix = "stand_") |>
  arrange(target_type, model)

make_performance_table <- function(
    scores, scores_summarised, 
    hor = 2, 
    cap = "Performance for two-week-ahead forecasts \\label{tab:scores}"
) {
  scores |>
    filter(scale %in% c("log", "natural")) |>
    summarise_scores(by = c("model", "target_type", "scale", "horizon")) |>
    select(model, target_type, scale, interval_score, coverage_50, coverage_90, horizon) |>
    pivot_wider(names_from = scale, values_from = interval_score) |>
    select(model, target_type, natural, log, coverage_50, coverage_90, horizon) |>
    inner_join(scores_summarised) |>
    mutate(across(c(where(is.numeric)), \(x) signif(x, digits = 2))) |>
    arrange(target_type, model) |>
    mutate(natural = label_fn(natural), 
           log = label_fn(log)) |>
    mutate(natural = paste0(natural, " (", label_fn(sd_stand_natural), ")"), 
           log = paste0(log, " (", label_fn(sd_stand_log), ")")) |>
    filter(horizon == hor) |>
    select(-mean_stand_log, -sd_stand_log, -mean_stand_natural, -sd_stand_natural, -horizon) |>
    kable(format = "latex", booktabs = TRUE, 
          align = c("l", "l", rep("c", 4)),
          caption = cap,
          linesep = c('', '', '', '\\addlinespace'),
          col.names = c("Model", "Target", "WIS", "WIS (log scale)", 
                        "Coverage 50%", "Coverage 90%")) |> 
    kable_styling(latex_options = c("scale_down", 
                                    "hold_position")) 
}

# Figure 1: Predictive performance across horizons (summary)
================================================================================

@@

ggsave("output/figures/performance.png", plot = p_performance, 
       width = 7, height = 6)
```

# Figure 2: Forecasts and corresponding WIS
================================================================================

@@

ggsave("output/figures/scores-and-forecasts.png", plot = p_scores_and_pred, 
       width = 9, height = 9)
```

# Figure 3: Comparison performance direct vs. Rt
================================================================================

ggsave("output/figures/comparison-direct-rt-individual.png", 
       plot = p_comparison_direct_rt_individual, 
       width = 7, height = 4)
```

================================================================================
#                                  Appendix                                    #
================================================================================

# Figure SI.1: Number of forecasters across the study period
================================================================================
- Figure SI.1
- numbers of forecasters used in the text

# Figure SI.2: Data and data revisions
================================================================================

Sources for data revisions: 
- https://github.com/CSSEGISandData/COVID-19/issues/6021
- https://coronavirus.data.gov.uk/details/whats-new 
- https://coronavirus.data.gov.uk/details/whats-new/record/965a7923-dacf-42ed-abb9-3a972dbf1835

# Figure SI.3 + SI.4: Screenshots from the forecasting app
================================================================================

# Figure SI.5: Performance of "experts" and "non-experts"
================================================================================
Separate ensemble of experts and non-experts

```{r}
create_expert_ensemble <- function(forecasts) {

@@

expert_ensemble$scores |>
   make_performance_table(
     scores_summarised = expert_scores_summarised,
     hor = 4,
     cap = "Performance for four-week-ahead forecasts of experts and non-experts. Standard deviations are given in brackets. \\label{tab:scores-experts-4}"
    ) %>%
  gsub("\\{table\\}", "\\{table\\*\\}", x = .) |>
  cat(file = 'manuscript/performance-table-experts-4.tex')

@@

@@

# Unused analyses
================================================================================

# Analysis on a comparable ensemble of Rt and direct forecasts
```{r}
# performance plot for an ensemble of the subset of human and direct forecasts
forecasts_comparable_ensemble <- forecasts_comparison |>
  group_by(target_type, horizon, method, quantile, true_value, forecast_date, target_end_date) |>
  summarise(prediction = median(prediction)) |>
  mutate(model = str_to_title(method)) |>
  as.data.table()

@@

# Analyses on revised data
```{r}
# forecasts_comparison_revised <- filter_comparison_rt_direct(forecasts_revised)
# scores_forecasts_comparison_revised <- score_and_add_log(forecasts_comparison_revised)

# p_comparison_direct_rt_individual_reivsed <- plot_rt_direct_individual(scores_forecasts_comparison_revised)
# 
# ggsave("output/figures/comparison-direct-rt-individual-revised.png", 
#        plot = p_comparison_direct_rt_individual, 
#        width = 7, height = 5)

# forecasts_comparable_ensemble_revised <- forecasts_comparison_revised |>
#   group_by(target_type, horizon, method, quantile, true_value, forecast_date, target_end_date) |>
#   summarise(prediction = median(prediction)) |>
#   mutate(model = str_to_title(method)) |>
#   as.data.table()
# 
# scores_comparable_ensemble_revised <- score_and_add_log(forecasts_comparable_ensemble_revised)
# 
# p_performance_comparison_revised <- plot_performance(
#   scores_comparable_ensemble_revised
# )
# 
# ggsave("output/figures/performance-rt-comparison-revised.png", plot = p_performance_comparison, 
#        width = 7, height = 8)
```

# Significance and bootstrap analysis

```{r}
scores |>
  filter(model %in% models_non_human) |>
  filter(horizon == 2) |>
  summarise_scores() |>
  select(model, forecast_date, type_and_scale, interval_score) |>
  pivot_wider(names_from = model, values_from = interval_score) |>
  group_by(type_and_scale) |>
  summarise(p1 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-direct`, paired = TRUE)$p.value, 
            p2 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-ensemble`, paired = TRUE)$p.value, 
            p3 = wilcox.test(`EuroCOVIDhub-ensemble`,`crowd-rt`, paired = TRUE)$p.value)
```

```{r}
# Only look at experts with direct forecasts
expert_ensemble_direct <- forecasts |>
  filter(model %in% models_direct) |>
  create_expert_ensemble()

expert_ensemble_direct$scores |>
  summarise_scores(c("model", "target_type", "horizon", "scale"))
   make_performance_table()

p_scores_expert_direct <- plot_performance(expert_ensemble_direct$scores)
ggsave("output/figures/performance-expert-direct.png", 
       plot = p_scores_expert_direct, 
       width = 7, height = 6)
```
